---
title: "Calculate statistics"
author: "Reto Stauffer"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
bibliography: annex.bib
link-citations: true
vignette: >
  %\VignetteIndexEntry{annex: Annex86 Data Analysis Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{Formula}
  %\VignetteKeywords{Annex86}
  %\VignettePackage{annex}
---

```{r setup, include = FALSE}
suppressPackageStartupMessages(library('zoo'))
suppressPackageStartupMessages(library('Formula'))

# Make a copy when building the article
file <- system.file("demos/demo_UIBK.xlsx", package = "annex", mustWork = TRUE)
file.copy(file, "demo_UIBK.xlsx", overwrite = TRUE)
```

# Introduction

Once the data is prepared, the analysis can be performed. This article
continues the example used in the article [Prepare data from XLSX](from_xlsx.html)
but works the very same way with the other examples included in the documentation, e.g.,
[Prepare data from textfiles](from_textfiles.html).

Please have a look at [this](from_xlsx.html) article for more information
on what the following lines of code do and/or to get the `"demo_UIBK.xlsx"` file
to follow this article.

```{r import_config_and_data}
library("readxl")
# Importing data and config object (meta information)
raw_df <- read_excel("demo_UIBK.xlsx", sheet = "measurements")
config <- read_excel("demo_UIBK.xlsx", sheet = "annex_configuration")
config <- subset(config, process == TRUE) # Custom subsetting

# Prepare data for annex()
library("annex")
prepared_df <- annex_prepare(raw_df, config, quiet = TRUE)
head(prepared_df, n = 3)
```

# Performing the analysis

Once the data set is prepared properly (note that `annex_prepare()`
is a convenience function, can also be done manually) the final object
can be prepared.

### Prepare `annex` object

`annex()` is the creator function which creates an object of class
`annex` (S3) providing a series of methods and functions to conduct the
final analysis. More information on object orientation system S3 can be found
e.g. [here](https://discdown.org/rprogramming/classes-and-methods.html#object-orientation)

The function expects a formula as input which describes how to process
the data. The three parts of the formula are:

* `<measurements to be processed> ~ <datetime> | <grouping variables>`
* The first part defines which variables (measurements) should be processed
* Part two is always `~ datetime`; the date and time information for the statistics
* Part three the grouping, typically `study + home + room`

```{r calling_annex}
annex_df <- annex(RH + T + CO2 ~ datetime | study + home + room,
                  data = prepared_df, tz = "Europe/Berlin")
head(annex_df)
class(annex_df)
```

A series of S3 methods exist for `annex` objects which might be extended
in the future.

* `summary()`
* `head()`/`tail()`
* `is.regular()`
* `plot()`

### Calculating statistics

Based on the object returned by `annex()` the analysis
can be performed by calling `annex_stats()`. The function
aggregates the data based on the formula provided above,
calculates a series of statistical properties, and returns
an object of class `annex_stats`.

```{r calling_annex_stats}
head(annex_df)
stats <- annex_stats(annex_df, format = "long")
head(stats)
```

By default, the argument format is set to `"wide"` which will return the
statistics in a wide format, i.e. all calculated values as columns while
`format = "long"` will create one line for every calculated value.
The chosen format (long or wide) does not matter for the further analysis but a particular
format may be convenient when processed manually.

## Next steps

After calculating the statistics, the following steps can be performed:

* [Visualizing the data (optional)](visualization.html)
* [Write annex stats to final file](write_and_validate.html#write)
* [Update meta data in final file](write_and_validate.html#meta)
* [Validate final file](write_and_validate.html#validate)


```{r, include = FALSE}
files <- c("demo_UIBK_config.csv", "demo_UIBK.xlsx", "final_UIBK.xlsx")
for (f in files) {
    if (file.exists(f)) file.remove(f)
}
```







