---
title: "Config feature"
author: "Reto Stauffer"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
bibliography: annex.bib
link-citations: true
vignette: >
  %\VignetteIndexEntry{annex: Annex86 Data Analysis Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{Formula}
  %\VignetteKeywords{Annex86}
  %\VignettePackage{annex}
---

## Introduction

To make the pre-processing of the data for the further analysis
more user-friendly and easier to automatize, the `annex` package
provides a configuration feature.

Technically is just a regular [`data.frame`](https://discdown.org/rprogramming/dataframes.html)
used as a lookup table to bring the data into the expected format which can come
from any source (e.g., from a text file, XLSX file, created via script, ...).

## Configuration content and purpose


```{r, include = FALSE}
file <- system.file("data/demo_Bedroom_config.TXT", package = "annex")
stopifnot(length(file) == 1, nchar(file) > 0, file.exists(file))
config <- read.table(file,
                     comment.char = "#", sep = "",
                     header = TRUE, na.strings = c("NA", "empty"))
```

The following output shows the first `n = 6` lines from the configuration object
used in the article [Analysis from textfiles](articles/from_textfiles.html);
the first six lines of a data.frame with $`r ncol(config)`$ variables (columns).

```{r}
head(config, n = 6)
```

This is used to tell the `annex` package which information needs to be processed,
where to find the important information, and how to translate (rename) the variables
to meet the `annex` standard format.

#### The five required columns

* `column`: The name of the column in the imported data set (e.g., from an
  [XLSX file](articles/from_xlsx.html) or [textfile](articles/from_textfiles.html)).
* `variable`: Defines the _new variable name_; one of the different variables (parameters)
  expected by the annex package.
* `study`: Name of the study the data belong to
* `home`: Name of the home (building) the data belong to
* `room`: Abbreviation of the type of room the data belong to

While `column`, `study` and `home` can be chosen freely by the user, only a series of
pre-defined values are available for `variable` and `room` (not case sensitive).

#### The required rows

There is one required row which must exist which is the one which defines
where the date and time information can be found. This row defines the
variable `"datetime"` and where it is located in the imported data set
(in this case a variable called `"X"`).

All other rows specify which column in the imported data set contains
and which information should be processed. Variables not listed in this
configuration (but available in the imported data set) will be ignored
in the processing steps (namely when calling `annex_prepare()`).


#### Purpose

The purpose of this object is to prepare an imported data set into
the object expected by the `annex` package for the analysis before
saving the data for later use.

It is used to define where the required information is stored and
used as a lookup table (translation table) to prepare/rename the
variables, to format the data according to the `annex` standard format
to be used for the further processing steps.


## Read/import configuration

The two articles
[Analysis from XLSX](articles/from_xlsx.html)
[Analysis from textfiles](from_textfiles.html)
show examples how to import (read) such a configuration from an XLSX file
or a text file. Any file which can be imported into a data.frame via _R_
can be used.

## Technical note

Note that the configuration is a convenience feature. Technically it can be
bypassed by preparing the data set outside annex before starting with the 
analysis.


