---
title: "Prepare data from XLSX"
author: "Reto Stauffer"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
bibliography: annex.bib
link-citations: true
vignette: >
  %\VignetteIndexEntry{annex: Annex86 Data Analysis Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{Formula}
  %\VignetteKeywords{Annex86}
  %\VignettePackage{annex}
---

```{r setup, include = FALSE}
suppressPackageStartupMessages(library('zoo'))
suppressPackageStartupMessages(library('Formula'))

# Make a copy when building the article
file <- system.file("demos/demo_UIBK.xlsx", package = "annex")
file.copy(file, "demo_UIBK.xlsx")
```

## Introduction

This article explains how to import and process data with the
`annex` package when all required data is available the XLSX format.

For this purpose, a demo data set called
`r xfun::embed_file("demo_UIBK.xlsx", text = "demo_UIBK.xlsx")` is used
(shipped with the package) which contains two sheets:

* **measurements**: The data (e.g., from a logger or pre-processed)
* **annex_configuration**: Configuration (see article [Config file](config.html))
Save the demo file and/or your xlsx data file in your R working directory.

Importing the data into R is up to the user, the `annex` package does not provide
import functionality. For XLSX files a multitude of R packages exist; this article
is making use of the [`readxl`](https://cran.r-project.org/package=readxl) package,
however, any other package which allows to read data from XLSX files and return
an object of class `data.frame` will work as well.

## Reading the data

The first step is to import both (i) the measurement data (stored on `raw_df`)
and (ii) the configuration (stored on `config`):

```{r import_config_and_data}
library("readxl")
raw_df <- read_excel("demo_UIBK.xlsx", sheet = "measurements")
config <- read_excel("demo_UIBK.xlsx", sheet = "annex_configuration")

# Class and dimension of the objects
c("raw_df" = is.data.frame(raw_df), "config" = is.data.frame(config))
cbind("raw_df" = dim(raw_df), "config" = dim(config))
```

Both objects are of class `data.frame` (tibble data frames to be precise)
with a dimension of $`r paste(dim(raw_df), collapse = " \\times ")`$ (`raw_df`)
and  $`r paste(dim(config), collapse = " \\times ")`$  (`config`) respectively.

The first few observations (rows) of the two objects look as follows:

```{r show_config_and_data}
head(raw_df[, 1:3], n = 3) # First three columns only
head(config, n = 3)
```

The object `raw_df` contains variables (columns) named
`r paste(sprintf("\"%s\"", names(raw_df)[1:3]), collapse = ", ")` which
are the original names from the XLSX sheet, the `config` object contains
the definition what the columns in `raw_df` contains and where they belong to.
For more details read the article about the [Config file](config.html).

The variable `config$process` is a custom column containing a logical
value whether or not the column should be used for further processing.
We would like to only process the rows where `config$proces == TRUE`.

```{r subsetting_config}
config <- subset(config, process == TRUE) # Remove all rows process = FALSE
```

As `config` is a pure `data.frame`, any subsetting can be used.



### Checking the config object

To check whether or not the `config` object is as expected by the `annex` package,
the function `annex_check_config()` can be used. In case problems would
be detected, an error will be thrown (see [Config file](config.html)).
Else, the function is silent as in this example:

```{r checking_config}
library("annex")
annex_check_config(config)
```

... no errors, the `config` object meets the `annex` requirements. Note that
this step is _not_ necessary as it will be performed automatically when
calling `annex_prepare()` but can be handy during development.


## Preparing data

Based on `raw_df` which still contains the raw data set read from
the XLSX file and the `config` object with the information about
the renaming (and which columns to be used), the data set for
further processing can be created.


```{r calling_annex_prepare}
prepared_df <- annex_prepare(raw_df, config, quiet = TRUE)
head(prepared_df)
```

`annex_prepare()` performs a series of tasks:

* Checking the `config` object (calls `annex_check_config()` internally).
  If the `config` object is valid,
* the variables (columns) in `raw_df` are renamed and checked to be of the
  correct class,
* informs the user if there are any columns in `raw_df` not included in
  `config` (just a note) and additional columns defined in `config` which
  do not occur in `raw_df`, and returns the modified (possibly subsetted) object,
* ensures that `datetime` is a proper datetime object (`POSIXt`).


The checks of missing/additional definitions in `config` are intended to inform
the user about possible misspecifications and will not result in an error.

## Ensure correct date and time (timezone)

<div class="alert alert-info" role="alert">
**Note:** `read_excel()` (from the
[`readxl`](https://cran.r-project.org/package=readxl)) package may
automartically detect and convert date and time information into `POSIXct`
(date and time object), however, as Excel has no notion of timezones, the information
returned will always be in UTC!
</div>

Investigate the class and the time zone attribute attached:

```{r}
class(prepared_df$datetime)         # Class of our first variable (Zeitraum ab)
attr(prepared_df$datetime, "tzone") # Time zone used
```

To enforce a different time zone (assuming the data has been recorded in a time zone
different than UTC) one need to acoerce the information back into a character without
time zone information (using `format()`) and back into `POSIXct` which allows to set
the correct time zone. For demonstration (to show the command) we convert the information
to 'US/Central' and back to 'UTC'.

```{r}
head(prepared_df$datetime, n = 2) # First three entries
# Coerce to 'US/Central'
prepared_df$datetime <- as.POSIXct(format(prepared_df$datetime), tz = "US/Central")
head(prepared_df$datetime, n = 2) # First three entries
attr(prepared_df$datetime, "tzone")
# Coerce back to 'UTC'
prepared_df$datetime <- as.POSIXct(format(prepared_df$datetime), tz = "UTC")
head(prepared_df$datetime, n = 2) # First three entries
attr(prepared_df$datetime, "tzone")
```

The time zone conversion can also be done on the `raw_df` object (if needed) but
needs to be addressed before calling `annex()` (see [Calculating statistics (analysis)](calculate_stats.html)) to ensure that this information is processed properly.

## Next steps

After performing the data preparation,
the following steps can be performed:

* [Calculating statistics (analysis)](calculate_stats.html)
* [Visualizing the data (optional)](visualization.html)
* [Write annex stats to final file](write_and_validate.html#write)
* [Update meta data in final file](write_and_validate.html#meta)
* [Validate final file](write_and_validate.html#validate)


```{r, include = FALSE}
files <- c("demo_UIBK_config.csv", "demo_UIBK.xlsx", "final_UIBK.xlsx")
for (f in files) {
    if (file.exists(f)) file.remove(f)
}
```







