---
title: "Lookup functions/definitions"
author: "Reto Stauffer"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
bibliography: annex.bib
link-citations: true
vignette: >
  %\VignetteIndexEntry{annex: Annex86 Data Analysis Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{Formula}
  %\VignetteKeywords{Annex86}
  %\VignettePackage{annex}
---

```{r setup, include = FALSE}
suppressPackageStartupMessages(library("DT"))
```


## Introduction

The package is using a series of 'lookup tables' and data sets to validate
the correctness the data.

This includes testing if the [config object](config.html) are in the
correct format or to [validate the XLSX file](write_and_validate.html).
In case problems are found, the error messages, warnings, and notes should include
hints on what has been going wrong.

A series of small helper functions exist which allow the user to get the definitions
used internally for these kind of tests.

## Variable definition

`annex_variable_definition()` returns a `data.frame` (or list) containing
the names of all allowed variables as well as some additional information.

```{r var_def}
require("annex")
var_def <- annex_variable_definition()
head(var_def, n = 3)
```

* `name`: Name of the variable (as used in column `variable` in the [config object](config.html))
* `required`: Whether or not additional variable information is required (used by `annex_validate()`)
* `lower`: Lower bound; observations below this threshold will be considered of bad quality. Used to calculate `quality_lower` by `annex_stats()`. `NA` if no bound defined.
* `upper`: Upper bound; observations above this threshold will be considered of bad quality. Used to calculate `quality_upper` by `annex_stats()`. `NA` if no bound defined.
* `allowed_units`: Comma separated list of allowed `unit`s as needed in the [config object](config.html) or `NA` if not defined. If set, the first entry is the 'annex standard unit'
    (see also [Unit conversion](unitconversion.html)).

The table below shows this table in a web-table form:

```{r var_def_table, echo = FALSE}
datatable(var_def, options = list(paging = TRUE, dom = "t"))
```

## Room definition

`annex_room_definition()` returns a `data.frame` containing
the names of all allowed room definitions. For each allowed room type
there is a base name (e.g., `"BED"`; Bedroom) which can be extended by adding
up to two numeric digits. This allows to define multiple bedrooms if needed
(see `examples_valid_labels` included below).
These are the labels as used in the [config object](config.html).

```{r room_def}
require("annex")
room_def <- annex_room_definition()
head(room_def, n = 3)
```

* `name`: Base name/base abbreviation of all available room types.
* `long_name`: Human readable description.
* `examples_valid_labels`: Examples of allowed names to label multiple rooms of the same
    type within one home.

The table below shows the full list in a web-table form:

```{r room_def_table, echo = FALSE}
datatable(room_def, options = list(paging = FALSE, dom = "t"))
```

## ISO3 country code

In the output XLSX file (sheet "META-home") the user needs to specify
the country (location) of each 'home' using the ISO3 (ISO3166 alpha-3) country codes.
A list of all available country codes is shipped with the `annex` package and can
be accessed using `annex_countries()`. This information is used when validating
the output XLSX (see [Write and validate](write_and_validate.html)).

```{r iso3_def}
require("annex")
iso3_def <- annex_countries()
head(iso3_def, n = 3)
```

The table below shows all `r nrow(iso3_def)` defined countries; searchable table.

```{r iso3_def_table, echo = FALSE}
datatable(iso3_def)
```


