---
title: "Prepare with time zones & units"
author: "Reto Stauffer"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
bibliography: annex.bib
link-citations: true
vignette: >
  %\VignetteIndexEntry{annex: Annex86 Data Analysis Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{Formula}
  %\VignetteKeywords{Annex86}
  %\VignettePackage{annex}
---

```{r setup, include = FALSE}
suppressPackageStartupMessages(library('zoo'))
suppressPackageStartupMessages(library('Formula'))
# Make a copy when building the article
file <- system.file("demos/demo_US.xlsx", package = "annex", mustWork = TRUE)
file.copy(file, "demo_US.xlsx")
```

## Introduction

This article once again shows how to import and prepare a data set to calculate
the `annex` statistics, focusing on the following two aspects:

* **Time zone:** How to deal with time zones
* **Unit conversion:** Automatically convert units of the observations.

For this demonstration a file called
`r xfun::embed_file("demo_US.xlsx", text = "demo_US.xlsx")` is used
which already contains two sheets, one which contains simulated observations,
and one containing the configuration (meta info) which allows to translate the
data in the 'observations sheet' into the object required by `annex()`.

Please also have a look at the two articles [Prepare data from XLSX](from_xlsx.html)
and [Prepare data from textfiles](from_textfiles.html) for more examples.

## Scenario **(1)**

Imagine you got observations from a logger which was recording the date and time
in the "US/Central" (CST) time and contains data from four different sensors:

* Kitchen room temperature (indoor); **Fahrenheit**
* Kitchen relative humidity (indoor); **no unit** (`0.0` to `1.0`)
* An exhaust temperature (heating system); **degrees Celsius**
* Ambient air pressure; **mmHg** (millimetre of mercury)

### Reading the data

Step one is to import the data; using the [`readxl`](https://cran.r-project.org/package=readxl)
package once again:

```{r}
library("readxl")
excel_sheets("demo_US.xlsx") # Checking sheet names
# Reading observations
raw_df <- read_excel("demo_US.xlsx", sheet = "airquality_data")
raw_df <- as.data.frame(raw_df) # Convert to simple data.frame
# Reading configuration
config <- read_excel("demo_US.xlsx", sheet = "meta_config")
config <- as.data.frame(config) # Convert to simple data.frame
```

A quick look at the first $3$ observations (rows) of the object `raw_df` shows
that we got a variable `loggertime` which seems to contain the date and time of 
the observation in the US format (`mm/dd/yyyy HH:MM:SS`) which we know represents
date and time in the "US/Central" time. However, when looking at the class of
the variable `raw_df$loggertime` we can see that it is of class character (`chr`; text)
wherefore we need to properly convert it before making use of it.

```{r}
head(raw_df, n = 3)
class(raw_df$loggertime)
```

In addition, this simulated data set contains a periods with missing data as well
as some periods where not all but single sensors did not report data which will
be shown later.


### Convert `loggertime`

To convert the date and time information into a proper datetime object (an object
of class `POSIXct`) including the proper time zone, we can call `as.POSIXct()`.

**Arguments:**

* `x`: The character vector `raw_df$loggertime` as we have it right now in our object.
* `format`: The format of the date and time information (see `?strftime` to get a list of all available options).
* `tz`: Time zone definition. Can be in the long format (e.g., `"US/Central"`) or short (e.g., `"CST"`).
  Call `OlsonNames()` to get a list of time zones (long format).

```{r}
# Convert and overwrite existing column
raw_df$loggertime <- as.POSIXct(x = raw_df$loggertime,
                                format = "%m/%d/%Y %H:%M:%S",
                                tz = "US/Central")
# Checking the variable
head(raw_df$loggertime)
attr(raw_df$loggertime, "tz")   # Shows time zone abbreviation
```

This step is important to ensure `annex()` later on knows how to interpret the
date and time information for the aggregation. Note that the time zone of the
raw data and the time zone when calculating the statistics _can be different_.

### The config object

Next, let us have a look at the `config` object.

```{r}
config
```

As described in detail in [Config file/objects](config.html) this object contains
information on how to interpret the raw data set. As shown, the unit column defines
that one temperature is recorded in Fahrenheit (`unit = "F"`), the relative humidity
without unit (`unit = "-"`), and the air pressure in millimeter of mercury (`unit = 'mmHg'`)
in the raw data.

This becomes important later when calling `annex_prepare()`.


### Data overview

To better understand what happens let us have a quick look at the data.

```{r}
summary(raw_df)
```

`summary()` returns a five number summary plus the arithmetic mean of each
of the variables in our `raw_df` object. As shown, the observations for the
kitchen temperature (second variable) has a mean of `r round(mean(raw_df[, 2]), 2)`
which corresponds to about 20 degrees Celsius. The average relative humidity is
`r round(mean(raw_df[, 3]), 3)` which corresponds to about `r round(mean(raw_df[, 3]) * 100)`
percent.

Alternatively we can quickly convert the `raw_df` into a time series object (now the
`loggertime` has been prepared correctly) to check the data visually which clearly
shows the gaps in the data:

```{r, fig = TRUE, fig.width = 10, fig.height = 8, out.width = "100%"}
library("zoo")
raw_zoo <- zoo(raw_df[, -1], raw_df[, 1])
plot(raw_zoo)
```


### Prepare the data

We can now use `raw_df` in combination with our `config` object or prepare the data
for the next step. In this step, the units will automatically be converted (for those
variables the conversion is defined).

```{r}
library("annex")
prepared_df <- annex_prepare(raw_df, config)
head(prepared_df)
summary(prepared_df)
```

The function also reshapes the data, however, one can see that the temperature
data (contains temperature data from the exhaust as well as the kitchen) now all
is in degrees Celsius; the relative humidity has been automatically converted to percent.


### Calculating statistics

Once the data set is prepared, `annex()` can be called to prepare the data set
for the final calculations. _Note_: So far, the date and time information is still
in the time zone of the logger. When calling `annex()` we can specify in which time
zone the statistics should be calculated.

**Correct solution:** According to our scenario we know that our study area is located
in the "US/Central" time zone. This is what we need to tell `annex()` to correctly
do the conversion.

```{r}
annex_df <- annex(T + RH + Pressure ~ datetime | study + home + room,
                  data = prepared_df, tz = "US/Central")
head(annex_df)
```

As shown, the first three observations (before 7 am local time) fall into the
period `tod = "23-07"` while the next ones fall into the period `tod = "07-23"`.


## Scenario **(2)**

For illustration, let us imagine the same situation where our study is located
in the US/Central time zone, however, imagine the `loggertime` is reported in UTC.

Thus, what is needed is to (i) convert the `loggertime` into UTC, but keep
the further analysis steps relative to the US/Central time zone (when calling `annex()`).

```{r}
# Re-reading the original data set; Store it on a new object raw_UTC
raw_UTC <- read_excel("demo_US.xlsx", sheet = "airquality_data")
# Convert loggertime; this time with 'tz = "UTC"'
raw_UTC$loggertime <- as.POSIXct(x = raw_UTC$loggertime,
                                 format = "%m/%d/%Y %H:%M:%S",
                                 tz = "UTC")
# Prepare the data relative to US/Central; same 'config' object as before
prepared_UTC <- annex_prepare(raw_UTC, config)
annex_UTC    <- annex(T + RH + Pressure ~ datetime | study + home + room,
                      data = prepared_UTC, tz = "US/Central")
```

When comparing the first six observations (rows) of our object
`annex_df` (all handled in US/Central time) and `annex_UTC` (raw data in UTC,
processing in US/Central) one can see the difference in the `tod` variable
(also affects `year` and `month`) ...

```{r}
head(annex_df)
head(annex_UTC)
```

## Next steps

After performing the data preparation,
the following steps can be performed:

* [Calculating statistics (analysis)](calculate_stats.html)
* [Visualizing the data (optional)](visualization.html)
* [Write annex stats to final file](write_and_validate.html#write)
* [Update meta data in final file](write_and_validate.html#meta)
* [Validate final file](write_and_validate.html#validate)



```{r, include = FALSE}
if (file.exists("demo_US.xlsx")) file.remove("demo_US.xlsx")
```






