---
title: "Visualization (plotting)"
author: "Reto Stauffer"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
bibliography: annex.bib
link-citations: true
vignette: >
  %\VignetteIndexEntry{annex: Annex86 Data Analysis Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{Formula}
  %\VignetteKeywords{Annex86}
  %\VignettePackage{annex}
---

```{r setup, include = FALSE}
suppressPackageStartupMessages(library('zoo'))
suppressPackageStartupMessages(library('Formula'))

# Make a copy when building the article
file <- system.file("data/demo_UIBK.xlsx", package = "annex")
file.copy(file, "demo_UIBK.xlsx")
```

## Introduction

The package comes with some simple default plots for `annex` objects
(as returned by `annex()`). The following demonstration is based on the
example as shown in the article [From XLSX](from_xlsx.html), using the
file `r xfun::embed_file("demo_UIBK.xlsx", text = "demo_UIBK.xlsx")`
(please read [From XLSX](from_xlsx.html) for details).

```{r prepare_dataset}
library("readxl")

# Reading measurement data
raw_df <- read_excel("demo_UIBK.xlsx", sheet = "measurements")

# Read and prepare config object
config <- read_excel("demo_UIBK.xlsx", sheet = "annex_configuration")
config <- subset(config, process == TRUE) # Remove all rows process = FALSE

# Prepare annex object
library("annex")
prepared_df <- annex_prepare(raw_df, config, quiet = TRUE)
annex_df    <- annex(RH + T + CO2 ~ datetime | study + home + room, data = prepared_df)
dim(annex_df)
class(annex_df)
```

### Plotting

Once prepared, we can call the default (base R) plotting methods:

```{r plotting_all, fig.width = 8, fig.height = 6, out.width = "100%"}
plot(annex_df)
```

Additional arguments `start` and `end` can be given (either objects
of class `POSIXt` or characters in the ISO format) to only plot
a specific time period.

```{r plotting_subset, fig.width = 8, fig.height = 6, out.width = "100%"}
plot(annex_df, start = "2011-10-03", end = "2011-10-04")
```


```{r, include = FALSE}
files <- c("demo_UIBK.xlsx")
for (f in files) {
    if (file.exists(f)) file.remove(f)
}
```

