---
title: "Visualization (plotting)"
author: "Reto Stauffer"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
bibliography: annex.bib
link-citations: true
vignette: >
  %\VignetteIndexEntry{annex: Annex86 Data Analysis Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{Formula}
  %\VignetteKeywords{Annex86}
  %\VignettePackage{annex}
---

```{r setup, include = FALSE}
suppressPackageStartupMessages(library('zoo'))
suppressPackageStartupMessages(library('Formula'))

# Make a copy when building the article
file <- system.file("data/demo_UIBK.xlsx", package = "annex")
file.copy(file, "demo_UIBK.xlsx")
```

## Introduction

The package comes with some simple default plots for `annex` objects
(as returned by `annex()`). The following demonstration is based on the
example as shown in the article [From XLSX](from_xlsx.html), using the
file `r xfun::embed_file("demo_UIBK.xlsx", text = "demo_UIBK.xlsx")`
(please read [From XLSX](from_xlsx.html) for details).

```{r prepare_dataset}
library("readxl")

# Reading measurement data
raw_df <- read_excel("demo_UIBK.xlsx", sheet = "measurements")

# Read and prepare config object
config <- read_excel("demo_UIBK.xlsx", sheet = "annex_configuration")
config <- subset(config, process == TRUE) # Remove all rows process = FALSE

# Prepare annex object
library("annex")
prepared_df <- annex_prepare(raw_df, config, quiet = TRUE)
annex_df    <- annex(RH + T + CO2 ~ datetime | study + home + room,
                     data = prepared_df, tz = "Europe/Berlin")
dim(annex_df)
class(annex_df)
```

### Plotting

Once prepared, we can call the default (base R) plotting methods:

```{r plotting_all, fig.width = 8, fig.height = 6, out.width = "100%"}
plot(annex_df)
```

Additional arguments `start` and `end` can be given (either objects
of class `POSIXt` or characters in the ISO format) to only plot
a specific time period.

```{r plotting_subset, fig.width = 8, fig.height = 6, out.width = "100%"}
plot(annex_df, start = "2011-10-03", end = "2011-10-04")
```


```{r, include = FALSE}
files <- c("demo_UIBK.xlsx")
for (f in files) {
    if (file.exists(f)) file.remove(f)
}
```

### Using `ggplot2`

The 'long form' of the stats can be handy for plotting with ggplot.

```{r}
stats <- annex_stats(annex_df)
```


To ensure we have the long form, `annex_stats_reshape()` can be called.
If it is already in the long format it will simply return itself, else
it will be reshaped from the wide format to the desired long format.

```{r}
stats <- annex_stats_reshape(stats, format = "long")
head(stats, n = 2)
```

**Empirical CDF:** The `stats` contains a series of quantiles
following the naming scheme `"^p[0-9\\.]+$"` (e.g., `p00`, `p02.5`, ...).
The code chunk below subsets the statistics to only get rows containing
these empirical quantiles, and appends an additional variable `p` with
the numeric value of the percentile as this is required for the
demo plot following afterwards.


```{r}
# Extracting rows with empirical quantiles
tmp    <- stats[grepl("^p[0-9\\.]+$", stats$stats), ]
# Extract numeric value for plotting
tmp$p  <- as.numeric(gsub("p", "", tmp$stats))
# Unique ID as an interaction of study, home, room, month, and tod
tmp$ID <- with(tmp, interaction(study, home, room, month, tod))
```

```{r viz_ggplot_T, fig = TRUE, fig.width = 10, fig.height = 8}
library("ggplot2")
ggplot(subset(tmp, variable == "T")) +
    geom_line(aes(x = p, y = value, col = home, group = ID)) +
    facet_wrap(~ room, scale = "free") +
    ggtitle("Empirical Cumulative Distribution Function of Temperature by Room Type")
```

```{r viz_ggplot_CO2, fig = TRUE, fig.width = 10, fig.height = 5}
ggplot(subset(tmp, variable == "CO2")) +
    geom_line(aes(x = p, y = value, col = interaction(month, tod, sep = " - "), group = ID)) +
    facet_wrap(~ room) +
    ggtitle("Empirical Cumulative Distribution Function of CO2 Concentration by Room Type")
```

```{r viz_ggplot_RH, fig = TRUE, fig.width = 10, fig.height = 8}
ggplot(subset(tmp, variable == "RH")) +
    geom_line(aes(x = p, y = value, col = room, group = ID)) +
    facet_wrap(~ home) +
    ggtitle("Empirical Cumulative Distribution Function of Relative Humidity by Home")
```



