---
title: "Write output"
author: "Reto Stauffer"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
bibliography: annex.bib
link-citations: true
vignette: >
  %\VignetteIndexEntry{annex: Annex86 Data Analysis Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{Formula}
  %\VignetteKeywords{Annex86}
  %\VignettePackage{annex}
---

```{r setup, include = FALSE}
suppressPackageStartupMessages(library('zoo'))
suppressPackageStartupMessages(library('Formula'))

# Make a copy when building the article
file <- system.file("data/demo_UIBK.xlsx", package = "annex")
file.copy(file, "demo_UIBK.xlsx")
```

## Introduction

This article demonstrates on how to write the processed data (annex
statistics) into the final pre-filled XLSX file and how to validate
the file format. The example is based on the demo data set
`r xfun::embed_file("demo_UIBK.xlsx", text = "demo_UIBK.xlsx")` again,
for details about the following code chunk please read the article
[From XLSX](from_xlsx.html).

```{r prepare_dataset}
library("readxl")

# Reading measurement data
raw_df <- read_excel("demo_UIBK.xlsx", sheet = "measurements")

# Read and prepare config object
config <- read_excel("demo_UIBK.xlsx", sheet = "annex_configuration")
config <- subset(config, process == TRUE) # Remove all rows process = FALSE

# Prepare annex object
library("annex")
prepared_df <- annex_prepare(raw_df, config, quiet = TRUE)
annex_df    <- annex(RH + T + CO2 ~ datetime | study + home + room, data = prepared_df)
stats       <- annex_stats(annex_df, format = "long")
head(stats)
```


## Write annex statistics to file {#write}

Once the statistics have been calculated, the data - as returned
by `annex_stats()` - can be written into the final ouput file.

```{r delete_output_for_docs_start, include = FALSE}
if (file.exists("final_UIBK.xlsx")) file.remove("final_UIBK.xlsx")
```

The function `annex_write_stats()` will create a new XLSX file
based on an `annex` template, store the calculated statistics,
and pre-fill a series of META sheets to take up the required
meta information. This information must be added manually by
the user.

```{r, warning = TRUE, error = TRUE}
annex_write_stats(stats, file = "final_UIBK.xlsx", user = 1234)
```

#### Where to get a user ID?

tbd [ask Gabi]

## Update meta information {#meta}

At this point, the XLSX file 
`r xfun::embed_file("final_UIBK.xlsx", text = "final_UIBK.xlsx")`
does contain all the statistics, but missing the meta information.

Where additional information is needed, the XLSX sheet contains
`<....>` (check any of the META sheets) to be updated.

At any time the file can be validated (see next section) to see
if everything is according to the requested standard of information
is still missing or invalid information has been entered.


## Validate final file {#validate}

The function `annex_validate()` performs a series of checks both
against the original template (included in the package; always
uses current version) as well as within the XLSX file provided
as input.

There are critical checks which will result in an exception (error;
stops execution) as well as checks resulting in notes for the user.
These notes will tell what's wrong or missing to allow the user
to make the appropriate changes, before validating the file again.

The function returns `TRUE` or `FALSE` (if it does not throw an
error), once `TRUE` is returned everything is OK with the file
and all required meta information has been entered properly.
At this stage, the file can be used for further processing or to
upload to the research repository (planned for the near future).

```{r, warning = TRUE, error = TRUE}
check <- annex_validate("final_UIBK.xlsx", user = 1234)
```

In his case no changes have been made to the file written by
`annex_write_stats()`. As no illegal changes have been made, no
critical errors are detected. However, a series of messages are
printed to inform that not everything has been filled in properly.

Simply open the file in the spreadsheet editing software of your
choice (MS Excel, Libreoffice, ...), make the required changes,
save the file, and repeat validation until the `annex_validate()`
is happy and returns `TRUE`.



```{r delete_output_for_docs_end, include = FALSE}
if (file.exists("final_UIBK.xlsx")) file.remove("final_UIBK.xlsx")
```




